---
title: "gdv - Fundamentals of data visualization"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

This report will cover my learnings with the fundamentals of data visualization, based on the data from the "Cross-Selling in Banking" challenge I completed last semester. I will take this opportunity to look at plots from the data visualization standpoint this time and improve on the results based on my new knowledge.

# LO 1: Visualization basics, chart types

In this first chapter, I will be looking at...

- different types of visualizations
- factors that play a role in deciding which visualization is a fit for what data 
  - considerations of visual variables 
  - perceptual theories
  - empirical evidence 
  - target group characteristics
  
## Density Plot
When we want to better understand the characteristics of continous data, a *density plot* can be really helpful.
This plot looks at the age distrubtion of the bank clients, by gender. It shows us the share of a given age (x-axis) and gender in the customer base.

I purposely made some adjustments to the standard plot, to align with what the age stands for logically.

- There are no customers younger than 10 years. However, every person was once younger, and I want to make clear for everybody looking at the plot that there are no customers under the age of 10. Therefore, the x-axis is fixed to start at 0.
- The standard labels on x would only show the age at 4 positions, which makes it difficult to read. To fix that, I changed the scale to be shown in an interval of 10 years.

```{r}
ggplot(df_client_prep, aes(x = age, color = gender)) +
  geom_density() +
  labs(title = "Age Distribution of the Bank Clients", x = "Age (years)", y = "Density", color = "Gender") +
  theme(legend.position = "bottom") +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 10)) +
  expand_limits(x = 0)
```

## Violin Plot

Another way of showing distrubtions is a violin plot. That allows a better distinction of the distributions using a discrete value.
Instead of lines on one plot, we can see the distrubtions side by side. I found this a really nice way of finding patterns in the data. Using comparisons, taking some time to look at it, differences stand out really well to the viewers eye.

In this specific case, we can see the correlation between loan status and loan size nicely. The colors used are also a good match to the meaning behind the status. Red is bad, green is good, pink is bad so far, annd yellow-green stands for good so far.

```{r plot-bank-products-2}
ggplot(df_loan_prep, aes(x = status, y = amount, color = status)) +
  geom_violin() +
  scale_y_continuous(labels = label_number(suffix = " Kč", scale = 1e+0)) + # find correct scale!!!
  labs(title = "Loan amount distribution by payment status", subtitle = "over all bank clients with a loan", x = "Status", y = "Amount") +
  theme(legend.position = "bottom") +
  scale_color_manual(values=loan_status_colors)
```